---
title: "Assignment2_Part1_VoiceInSchizophrenia"
author: "Riccardo Fusaroli"
date: "July 17, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 2 - Part 1 - Assessing voice in schizophrenia

0 = control 
1 = schizophrenia
study 1 = 100, study 2 = 200 etc. 
trial = videos they have described

Schizophrenia has been associated with "inappropriate" voice, sometimes monotone, sometimes croaky. A few studies indicate that pitch might be an index of schizophrenia. However, an ongoing meta-analysis of the literature (which you will have a go at in the last assignment) indicates that pitch mean and standard deviation are only weak indicators of diagnosis. Can we do better with our new fancy complex skills?

The corpus you are asked to analyse is a set of voice recordings from people with schizophrenia (just after first diagnosis) and 1-1 matched controls (on gender, age, education). Each participant watched 10 videos of triangles moving across the screen and had to describe them (so you have circa 10 recordings per person). I have already extracted the pitch once every 10 milliseconds and you will have to use this data to assess differences in the voice.

N.B. Question to be answered via email to Celine: can you characterize voice in schizophrenia as acoustically different? Report the methods you used to answer this question and the results from the analyses. Add a couple of lines trying to interpret the results (make sense of the difference). E.g. People with schizophrenia tend to have high-pitched voice, and present bigger swings in their prosody than controls. Add a couple of lines describing limitations of the data/analyses if any is relevant.

N.B. There are looots of files to be dealt with. Probably too many for your computer. This is a challenge for you. Some (complementary) possible strategies: You can select a subset of files only (and you have to justify your choice). You can learn how to use the apply() or map() functions. You can coordinate with classmates.

```{r}

#Choose directory
setwd("~/Desktop/Experimental methods 3/Assignment 3")

#Load files/data
articulation = read.delim("Articulation.txt", sep=",")
demo_data = read.delim("DemoData.txt")
T2 = read.delim("Study1D0S103T2_f0.txt")

#Load libraries
library(pastecs)
library(modelr)
library(Metrics)
library(crqa)
library(ggplot2)
library(stringr)
library(readr)

```

1. In the course of this assignment you have to first select one datafile and figure out how to:

- Extract "standard" descriptors of pitch: Mean, standard deviation, range
- Extract less "standard" descriptors of pitch you can think of (e.g. median, iqr, mean absoluted deviation, coefficient of variation)
- Extract "complex" descriptors: recurrence quantification analysis
```{r}

stat.desc(T2$f0)

IQR(T2$f0)

Metrics::rmse(T2, T2$f0)

#Recurrence quantification analysis
#Finding the optimal parameters: run on all pairs: 
par = list(lgM =  50, steps = seq(1, 6, 1),  radiusspan = 100,  radiussample = 40, normalize = 0,  rescale = 0,  mindiagline = 2,  minvertline = 2,  tw = 0,  whiteline = FALSE,  recpt = FALSE,  fnnpercent = 10,  typeami = "mindip") 

ans = optimizeParam(T2$f0, T2$f0, par, min.rec= 3.5, max.rec= 4.5)

#Choose a common value for delay, emb.dim and radius 
Results=crqa (T2$f0, T2$f0, delay=ans$delay, embed=ans$emddim, radius=ans$radius,normalize=0,rescale=0,mindiagline = 2,minvertline = 2) 

#Represent the plot: 
RP=Results$RP 
RP = matrix(as.numeric(RP), nrow = ncol(RP))  
cols = c("white","blue4") 
image(RP, xlab = "", ylab = "", col = cols) 

#Explore the lags of coordination: 
Profile=drpdfromts(T2$f0, T2$f0, datatype = 'continuous',ws=50,radius=ans$radius) 
timecourse = round( seq(-5000,5000,100)/1000, digit = 1) 
maxlag = Profile$maxlag/1000 
profile = Profile$profile*100 
Prof=data.frame(profile) 
ggplot(Prof, aes(timecourse,profile))+geom_line()+ geom_vline(xintercept = timecourse[maxlag], colour='red')

```

IQR how likely is the pitch to vary = 16.65hz. 
The pitch can vary 59.5 in both ways from the mean.

mean = 211

range = 132.4

standard deviation = 15.7 #Our data is not normally distributed which this function relies on, therefore we cannot use this descrptive stats. Use IQR

median = 208

mean absoluted deviation = 146.3

coefficient of variantion = 0.07 


2. Second you will have to turn the code into a function and loop through all the files (or even better use apply/sapply/lapply)
- Remember to extract the relevant information from the file names (Participant, Diagnosis, Trial, Study)

```{r}

# pitch_files = list.files(pattern="*f0.txt")
# files = lapply(pitch_files, read.delim) 
# 
# Subject = NULL
# Diagnosis= NULL
# Trial = NULL
# Study = NULL
# Delay = NULL
# Radius = NULL
# Embed = NULL
# n = 1
# 
# for (i in pitch_files) {
#   print(i)
#   file = read_delim(i, delim = "\t")
#   Study[n]= substr(i, 6, 6 )
#   Diagnosis[n]=substr(i, 8, 8)
#   Subject[n] = substr(i,10,12 ) #alternatively: substr(i, start = 10, stop=12)
#   Trial[n]= substr(i, 14, 14)
#     
#   n = n+1    
# }
# 
# 
# 
# delay=NULL
# radius=NULL
# embed=NULL
# 
# for (i in files){
#   crqa (T2$f0, T2$f0, delay=ans$delay, embed=ans$emddim, radius=ans$radius,normalize=0,rescale=0,mindiagline = 2,minvertline = 2) 
#   
#   par = list(lgM =  50, steps = seq(1, 6, 1),  radiusspan = 100,
#   radiussample = 40, normalize = 0,  rescale = 0,  mindiagline = 2, 
#   minvertline = 2,  tw = 0,  whiteline = FALSE,  recpt = FALSE, 
#   fnnpercent = 10,  typeami = "mindip") 
#   
#   ans = optimizeParam(T2$f0, T2$f0, par, min.rec= 3.5, max.rec= 4.5)  
#   
#   delay[n]= ans$delay
#   raduis[n]= ans$radius
#   embed[n]= ans$emddim
#     
#   n = n+1
# 
# }


# Beregn en værdi delay, embed og radius f.eks. mean


# Create a table with all the data combined(rbind)
#all_data = do.call(rbind, lapply(gender_data, function(x) read.csv(x, stringsAsFactors = FALSE)))
  
#data frame  
#face_data[[i]] = read.csv(temp[i])}
#face_data=ldply(face_data,rbind)


#FIRST LOOP WHICH WORKS
temp = list.files(pattern="*f0.txt")
myfiles = lapply(temp, read.delim)

#empty varaibles that can be filled
SUBJ <- NULL
study <-  NULL
trial <-  NULL
diagnosis <-  NULL
delay = NULL
embed = NULL
radius = NULL
mad = NULL
coef_var = NULL
mean <- NULL
sd <- NULL
range <- NULL
median <- NULL
IQR = NULL

# start from one
n = 1

#file = readr::read_delim(i, delim = "\t")

for (i in temp){
  print(i)
  df = readr::read_delim(i, delim = "\t")
  df = df$f0
  #with ID = str_extract(i,"S+\\d+") regular expressions 
  #participant[n]= str_extract(ID, "\\d+")
  study[n] = substr(i, start = 6, stop = 6)
  diagnosis[n]= substr(i, start = 8, stop = 8)
  SUBJ[n]= substr(i, start = 10, stop = 12)
  trial[n]=substr(i, start = 14, stop = 14)
  mean[n] <- mean(df, na.rm = T)
  sd[n] <- sd(df, na.rm =T)
  range[n] <- range(df, na.rm = T)[2]-range(df, na.rm = T)[1]
  median[n] <- median(df,na.rm = T)
  IQR[n] <- IQR(df, na.rm = T)
  mad[n] <- mad(df,center = mean(df), na.rm = T)
  coef_var[n] <- sd(df)/mean(df)*100
  
  par = list(lgM =  15, steps = seq(1, 6, 1),  radiusspan = 100, radiussample = 40, normalize = 0,  rescale = 0,  mindiagline = 2,  minvertline = 2,  tw = 0,  whiteline = FALSE,  recpt = FALSE,  fnnpercent = 10,  typeami = "mindip")
  #widen the range of the recurrence (3-5 instead)
  ans = try(optimizeParam(df, df, par, min.rec = 3, max.rec = 5))
  if(length(ans) < 2) {
    delay[n] = NA
    embed[n] = NA
    radius[n] = NA
  }
  else {
    delay[n]=ans$delay
    embed[n]=ans$emddim
    radius[n]=ans$radius
  }
  n=n+1 #making sure that the next is saved in the new row
}

delaym= round(median(delay, na.rm=T) - sd(delay, na.rm = T))
embedm= round(median(embed, na.rm=T) - sd(embed, na.rm = T))
radiusm=round(median(radius, na.rm=T))

    
#maximize the comparability

Data = data.frame(SUBJ, study, trial, diagnosis,  mean, sd, range, median, delay, embed, radius, mad, IQR, coef_var)



# SECOND LOOP WHICH WORKS
temp = list.files(pattern="*f0.txt")
myfiles = lapply(temp, read.delim)

PR = NULL
DET <- NULL
maxL <- NULL
L <- NULL
ENTR <- NULL
LAM <- NULL
TT <- NULL

n = 1

for (i in temp){
  print(i)
  df = readr::read_delim(i, delim = "\t")
  df = df$f0
  results = try(crqa(df,df, delay=delaym, embed=embedm, radius=3,normalize=0,rescale=0,mindiagline = 2,minvertline = 2))
  # Radius is set to 3 rather than the median of the radius due to the fact that the optimal recurrence percentage is 5%-10%
  
  if (length(results) > 1) {
    PR[n] = results$RR
    DET[n] <- results$DET
    maxL[n] <- results$maxL
    L[n] <- results$L
    ENTR[n] <- results$ENTR
    LAM[n] <- results$LAM
    TT[n] <- results$TT
  }
  else {
    PR[n] = NA
    DET[n] <- NA
    maxL[n] <- NA
    L[n] <- NA
    ENTR[n] <- NA
    LAM[n] <- NA
    TT[n] <- NA
  }
  n = n+1
}

data2 = data.frame(SUBJ, study, trial, diagnosis,  mean, sd, range, median, delay, embed, radius, mad, IQR, coef_var, PR, DET, maxL, L, ENTR, LAM, TT)


write.csv(final_df, file = "A3_P1_data.csv")

```

3. Make one model per acoustic feature and test whether you can observe significant difference due to Diagnosis. Tip: Which other fixed factors should you control for (that is, include in the model)? Which random ones?
- Bonus points: cross-validate the model and report the betas and standard errors from all rounds to get an idea of how robust the estimates are. 
3a. Is study a significant predictor in these models? What should you infer from this? Does study interact with diagnosis? What should you infer from this?

```{r}
library(lmerTest)
library(lme4)

data2$trial=as.numeric(data2$trial)
data2$diagnosis=as.numeric(data2$diagnosis)

model1 = lmer(mean ~ 1 + diagnosis + trial + (1+trial + diagnosis|SUBJ), REML=F, data = data2)
summary(model1)

```

4. Bonus Question: Compare effect size of diagnosis across the different measures. Which measure seems most sensitive?
- Tip: to compare across measures you need to put all of them on the same scale, that is, you need to "standardize" them (z-score)

```{r}

```

5. Bonus question. In the Clinical Info file you have additional information about the participants. Which additional parameters (e.g. age, gender) should we control for? Report the effects.

```{r}

```

6. Write a paragraph reporting methods and results

```{r}
We expect that participants with schizophrenia have a higher ptich and therefore must the mean of the pitch range also be higher compared to controls. 
```

[Next assignment: can we use these measures to build a tool that diagnoses people from voice only?]

## N.B. Remember to save the acoustic features of voice in a separate file, so to be able to load them next time